---
title: "Presidential birthplaces"
author: "Erica Pehrsson"
date: "February 28, 2019"
output: html_document
---

Map of the birth states of US presidents, colored by party. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load library, include=FALSE}
library(leaflet)
```

```{r birthplaces data frame, echo=FALSE}
# Load states
states = read.table("states.txt",sep='\t',header=TRUE)

# Load birthplaces
birthplaces = read.table("birthplaces.txt",sep='\t',header=TRUE)
birthplaces$State.of.birth = factor(birthplaces$State.of.birth,levels=levels(states$State))
birthplaces[which(birthplaces$President == "Andrew Jackson"),]$State.of.birth = "South Carolina"

# Load parties
parties = read.table("parties.txt",sep='\t',header=TRUE)
colnames(parties)[1] = "Number"
parties$President = factor(parties$President,levels=levels(birthplaces$President))
parties[is.na(parties)] = "William H. Taft"

# Merge
presidents = merge(parties,unique(birthplaces),by="President")
presidents = merge(presidents,states,by.x="State.of.birth",by.y="State")
```

```{r make map, echo=FALSE}
pal <- colorFactor(c("blue","purple","yellow","blue","red","red","grey","green","green"), presidents$Party)

map <- leaflet(data=presidents) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~Longitude, lat = ~Latitude, color = ~pal(Party), label= ~President, clusterOptions = markerClusterOptions())

map
```

Party legend: Democratic (blue; see 1), Republican (red; see 2), Democratic-Republican (purple), Whig (green; see 3), Federalist (yellow), Unaffiliated (grey).

1. Including Andrew Johnson, who switched from National Union
2. Including Abraham Lincoln, who switched to National Union
3. Including John Tyler, who switched to Unaffiliated

References:

* List of presidents: https://en.wikipedia.org/wiki/List_of_Presidents_of_the_United_States 
* List of presidential birthplaces: https://en.wikipedia.org/wiki/List_of_Presidents_of_the_United_States_by_home_state
* Latitude and longitude of states: https://inkplant.com/code/state-latitudes-longitudes